\large \textbf{Title:} Computational models reveal how cleaner fish adjust decisions in a biological market

### 2022-08-08 manuscript_BE/manuscript_BE.Rmd
Edits in the source code to fit parameters using PAA

\begin{equation}
\label{eq:delta}
\delta_t = R_t -  V_t(S_t) + \gamma V_t(S_{t+1}) ,
\end{equation}

@@

To capture with the model the relationship between the ecological 
variables and cleaner fish preferences for visitors, we needed to scale 
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures client visitation patterns. 
This is because, in the model, relative abundances of clients 
define not only the probability of residents and visitors 
but also how often the cleaning station is empty (e.g. there are no 
clients to be cleaned). The frequency with which clients visit the station 
is another variable influencing the station occupancy, which in nature 
may vary among different client species 
depending on their ectoparasite loads. We do not have field estimates 
for species-specific parasite loads, especially not as a function of the site. 
In order to control for these aspects, we computed a measure of relative 
cleaner fish abundance for each reef site relative to 
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. This scaling 
constant is hence meant to capture variation in the market conditions driven 
partly by cleaner fish abundance. We fitted the value of the scaling constant as 
part of the statistical inference. As for the visitor and resident 
abundances, we computed a relative measure with respect to the 
total client abundance, and weighted that by the rescaled relative absence 
of cleaners ($1-\text{relative cleaner abundance}$). Thus, all three measures 
of relative abundance sum up to 1, and can be used in the model as a proxy 
for the probability of having different options in the cleaning station. 
Note that we introduced the scaling constant to control for variation that 
is not captured by the model; its parameter distribution does not offer 
biological insights.

Once we calculated the relative abundances, we obtained predictions from 
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the 
computational model once for every reef site using as input the ecological 
explanatory variables. The model outputs the probability of choosing the visitor
for each location $p_i$, where $i$ is the index for the 12 locations. Assuming a 
binomial distribution, the probability that each of the cleaner fish 20 choices 
in the ephemeral reward task was generated by the model 
is given by $\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}$;
where $k_{ij}$ is the number of times that cleaner fish $j$ in reef site $i$ 
chose the visitor over the resident; and $n=20$ (due to our choice of using 20 
choices per cleaner fish). By taking the natural logarithm and summing over all 
the cleaner fish and reef sites, we obtained the log-likelihood of the data 
given the model and parameters. We then proposed a new set of parameters drawn 
from a uniform distribution centred around the old parameter set. The amplitude 
of the uniform distribution used for each parameter can be found in Table 
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood 
with the new parameter set. We then used the ratio of the two likelihoods to 
choose which parameter set to keep. New parameter sets with a higher likelihood 
than the old set replaced old ones, and those with a lower likelihood replaced 
current ones  with a probability equal to the log-likelihood ratio. 
Given that we only used the likelihoods in the decision, we used an 
uninformative prior. Once we decided whether the new parameter set 
would replace the old one, we ran the model again to sample the likelihood 
distribution of the parameter set. We then started 
the cycle again by proposing a new set of parameters and repeated
the process for $1e^5$ steps. We ran 5 independent chains, 
discarded the first $1000$ samples of each chain as burn in, 
and after that, we kept 1 in every 100 samples to avoid autocorrelation. 
The collection of parameters sets kept in 
all chains approximates the posterior distribution. We coded the model 
as well as the fitting algorithm in *c++*, the diagnostics and visualization 
in R [@rcoreteam_Language_2021]. All codes are accessible at 
[![DOI](https://zenodo.org/badge/440585701.svg)](https://zenodo.org/badge/latestdoi/440585701).

@@

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism 
cleaners use to account for the future effects of their actions;
and thus to solve the ephemeral reward task. In the estimation of the 
parameter values of the full model, which includes both chaining and
penalty, the bulk of the marginal posterior distribution of $\eta$ which 
controls the strength of penalty is around 0 (Fig. \ref{fig:post}). 
As for $\gamma$, controlling chaining, the 95% confidence 
intervals also includes zero, but the mode of the posterior 
is around 0.5 (Fig. \ref{fig:post}, a). In the chaining model, 
where $\eta$ is set to zero, the distribution of $\gamma$ shifted 
to higher values, zero is no longer part of the 95% credible interval
of the parameter (Fig. \ref{fig:post}, c). In contrast, when we 
look at the model with only penalty, the posterior distribution 
of $\eta$ is still centred around zero (Fig. \ref{fig:post}, d). 
Thus, the analysis of the estimates of individual parameter 
values in the three models only supports a strong 
effect of chaining. Furthermore, the comparison of the models` fit 
favours the chaining model. In panel e of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before. 
Note, $pseudo-R^2$ can have negative values, which is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though the peak of the three
$pseudo-R^2$ distributions were not very different, the model with only
chaining produced a distribution of $pseudo-R^2$ where more values
were positive (to the right of the black line in Fig. \ref{fig:post} e). This
shows that accounting for variation in the parameter estimates the
model with chaining gives a better fit to the data, despite
having one parameter less than the full model. We have not shown 
here the marginal posterior distributions of the scaling constant, 
given that they do not bring biological insight. 
Their visualizations can be found in the supplementary 
material (Fig. \ref{fig:scaConst}), as well as the diagnostics of the MCMCs
(Figs. \ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagNeg}).

@@

The main reason for chaining and penalty to give different
predictions is the way that cleaner fish relative abundance influences the
preference for the visitor clients. Visitor leaving probability has a
similar positive effect on the probability of choosing the visitor
clients on all three models, they all predict an increase preference 
for visitors as the visitor probability increases  (Fig. \ref{fig:pred}).
In contrast, cleaner fish relative abundance has a different effect in the
model with only chaining, compared to the other two models,
full model and penalty. In the model with only chaining, only intermediate 
cleaner fish abundance triggers a preference for the visitor 
clients (Fig. \ref{fig:pred} c). In both models with penalty, 
both intermediate and low cleaner abundance triggered a preference 
for the visitor (Fig. \ref{fig:pred} a,e). Note, however, we calculated 
preferences shown in figure \ref{fig:pred} left panels by using only the mode 
of the posterior distributions, and by holding constant the balance between 
resident and visitors´ abundances. Panels on the right, show how close 
predictions are from the observed data, allowing the balance between 
client types to vary and using a set of samples from the posterior 
distribution.

>>>>

\begin{equation}
\delta_t = R_t -  V_t(S_t) + \gamma V_t(S_{t+1}) ,
\label{eq:delta}
\end{equation}

@@

To capture with the model the relationship between the ecological 
variables and cleaner fish preferences for visitors, we needed to scale 
the absolute population densities of cleaner fish from the empirical data
to a measure of relative abundances that captures client visitation patterns. 
This is because, in the model, relative abundances of clients 
define not only the probability of residents and visitors 
but also how often the cleaning station is empty (e.g. there are no 
clients to be cleaned). The frequency with which clients visit the station 
is another variable influencing the station occupancy, which in nature 
may vary among different client species 
depending on their ectoparasite loads. We do not have field estimates 
for species-specific parasite loads, especially not as a function of the site. 
In order to control for these aspects, we computed a measure of relative 
cleaner fish abundance for each reef site relative to 
absolute abundances and multiplied it
by a scaling constant that changes the range of the variable. Specifically,
cleaner fish relative abundance is equal 
to $\frac{\epsilon C_{abs}}{\epsilon C_{abs}+R_{abs}+V_{abs}}$. 
Where $\epsilon$ is the scaling constant; $C_{abs}$ 
is the cleaner fish absolute density; $R_{abs}$ is the residents absolute 
density and $V_{abs}$ is the visitor client absolute density. This scaling 
constant is meant to capture variation in the market conditions driven 
partly by cleaner fish abundance. We fitted the value of the scaling constant
($\epsilon$) as part of the statistical inference. 
As for the visitor and resident abundances, we computed a 
relative measure with respect to the 
total client abundance, and weighted that by the re-scaled relative absence 
of cleaners ($1-\text{relative cleaner abundance}$). Thus, all three measures 
of relative abundance sum up to 1, and can be used in the model as a proxy 
for the probability of having different options in the cleaning station. 
Note that we introduced the scaling constant to control for variation that 
is not captured by the model; its parameter distribution does not offer 
biological insights.

Once we calculated the relative abundances, we obtained predictions from 
the model for each one of the locations and ran the Markov Chain. We started the
chain with random values for the parameters of interest, then ran the 
computational model once for every reef site using as input the ecological 
explanatory variables. The model outputs the probability of choosing the visitor
for each location $p_i$, where $i$ is the index for the 12 locations. Assuming a 
binomial distribution, the probability that each of the cleaner fish 20 choices 
in the ephemeral reward task was generated by the model 
is given by $\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}$;
where $k_{ij}$ is the number of times that cleaner fish $j$ in reef site $i$ 
chose the visitor over the resident; and $n=20$ (due to our choice of using 20 
choices per cleaner fish). By taking the natural logarithm and summing over all 
the cleaner fish and reef sites, we obtained the log-likelihood of the data 
given the model and parameters. We then proposed a new set of parameters drawn 
from a uniform distribution centered around the old parameter set. The amplitude 
of the uniform distribution used for each parameter can be found in Table 
\ref{tab:param}. Subsequently, we ran the model and calculated the likelihood 
with the new parameter set. We then used the ratio of the two likelihoods to 
choose which parameter set to keep. New parameter sets with a higher likelihood 
than the old set replaced old ones, and those with a lower likelihood replaced 
current ones  with a probability equal to the log-likelihood ratio. 
Given that we only used the likelihoods in the decision, we used an 
uninformative prior. Once we decided whether the new parameter set 
would replace the old one, we ran the model again to sample the likelihood 
distribution of the parameter set. We then started 
the cycle again by proposing a new set of parameters and repeated
the process for $1e^5$ steps. We ran 5 independent chains, 
discarded the first $1000$ samples of each chain as burn in, 
and after that, we kept 1 in every 100 samples to avoid autocorrelation. 
The collection of parameters sets kept in 
all chains approximates the posterior distribution. We coded the model 
as well as the fitting algorithm in *c++*, the diagnostics and visualization 
in R [@rcoreteam_Language_2021]. All codes are accessible at 
\url{https://zenodo.org/badge/latestdoi/440585701}
<!-- [![zenodo.6338073.svg)](https://zenodo.org/badge/latestdoi/440585701). -->

@@

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism 
cleaners use to account for the future effects of their actions;
and thus to solve the ephemeral reward task. In the estimation of the 
parameter values of the full model, which includes both chaining and
penalty, the bulk of the marginal posterior distribution of $\eta$ which 
controls the strength of penalty is around 0 (Fig. \ref{fig:post}). 
As for $\gamma$, controlling chaining, the 95% confidence 
intervals also includes zero, but the mode of the posterior 
is around 0.5 (Fig. \ref{fig:post}, a). In the chaining model, 
where $\eta$ is set to zero, the distribution of $\gamma$ shifted 
to higher values, zero is no longer part of the 95% credible interval
of the parameter (Fig. \ref{fig:post}, c). In contrast, when we 
look at the model with only penalty, the posterior distribution 
of $\eta$ is still centered around zero (Fig. \ref{fig:post}, d). 
Thus, the analysis of the estimates of individual parameter 
values in the three models only supports a strong 
effect of chaining. Furthermore, the comparison of the models' fit 
favors the chaining model. In panel e of figure \ref{fig:post} 
we show the distribution of $pseudo-R^2$ calculated using 
samples from the posterior distributions shown before. 
Note, $pseudo-R^2$ can have negative values, which is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though the peak of the three
$pseudo-R^2$ distributions were not very different, the model with only
chaining produced a distribution of $pseudo-R^2$ where more values
were positive (to the right of the black line in Fig. \ref{fig:post} e). This
shows that accounting for variation in the parameter estimates the
model with chaining gives a better fit to the data, despite
having one parameter less than the full model. We have not shown 
here the marginal posterior distributions of the scaling constant, 
given that they do not bring biological insight. 
Their visualizations can be found in the supplementary 
material (Fig. \ref{fig:scaConst}), as well as the diagnostics of the MCMCs
(Figs. \ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagNeg}).

@@

Low cleaner fish relative abundance yields different predictions 
from a model with only chaining and with penalty.
The model with only chaining predicts a low preference fpr the visitor when
cleaner fish relative abundance is low (Fig. \ref{fig:pred} c). In contrast,
the models including penalty (with and without chaining) predict high 
preference for the visitor option (Fig. \ref{fig:pred} a and e).
For intermediate and high cleaner relative abundances the predictions 
are similar for all three models. Intermediate relative abundance 
trigger a high preference for the visitor; while low abundances trigger
low preference for the visitor (Fig. \ref{fig:pred}a,b,e). 
Note, however, we calculated 
preferences shown in figure \ref{fig:pred} left panels by using only the model 
of the posterior distributions, and by holding constant the balance between 
resident and visitors' abundances. Panels on the right, show how close 
predictions are from the observed data, allowing the balance between 
client types to vary and using a set of samples from the posterior 
distribution.

Visitor leaving probability has an overall 
positive effect on the probability of choosing the visitor
clients on all three models. All three models predict an increase preference 
for visitors as the visitor probability increases  (Fig. \ref{fig:pred}).

### 2022-08-08 manuscript_BE/manuscript_BE.tex
Edits in the source code to fit parameters using PAA

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{graphicx}
\usepackage{booktabs}
\linespread{2}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}\doublespacing
\newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

@@

To capture with the model the relationship between the ecological
variables and cleaner fish preferences for visitors, we needed to scale
the absolute population densities of cleaner fish from the empirical
data to a measure of relative abundances that captures client visitation
patterns. This is because, in the model, relative abundances of clients
define not only the probability of residents and visitors but also how
often the cleaning station is empty (e.g.~there are no clients to be
cleaned). The frequency with which clients visit the station is another
variable influencing the station occupancy, which in nature may vary
among different client species depending on their ectoparasite loads. We
do not have field estimates for species-specific parasite loads,
especially not as a function of the site. In order to control for these
aspects, we computed a measure of relative cleaner fish abundance for
each reef site relative to absolute abundances and multiplied it by a
scaling constant that changes the range of the variable. This scaling
constant is hence meant to capture variation in the market conditions
driven partly by cleaner fish abundance. We fitted the value of the
scaling constant as part of the statistical inference. As for the
visitor and resident abundances, we computed a relative measure with
respect to the total client abundance, and weighted that by the rescaled
relative absence of cleaners (\(1-\text{relative cleaner abundance}\)).
Thus, all three measures of relative abundance sum up to 1, and can be
used in the model as a proxy for the probability of having different
options in the cleaning station. Note that we introduced the scaling
constant to control for variation that is not captured by the model; its
parameter distribution does not offer biological insights.

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We
started the chain with random values for the parameters of interest,
then ran the computational model once for every reef site using as input
the ecological explanatory variables. The model outputs the probability
of choosing the visitor for each location \(p_i\), where \(i\) is the
index for the 12 locations. Assuming a binomial distribution, the
probability that each of the cleaner fish 20 choices in the ephemeral
reward task was generated by the model is given by
\(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\); where \(k_{ij}\) is
the number of times that cleaner fish \(j\) in reef site \(i\) chose the
visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing
over all the cleaner fish and reef sites, we obtained the log-likelihood
of the data given the model and parameters. We then proposed a new set
of parameters drawn from a uniform distribution centred around the old
parameter set. The amplitude of the uniform distribution used for each
parameter can be found in Table \ref{tab:param}. Subsequently, we ran
the model and calculated the likelihood with the new parameter set. We
then used the ratio of the two likelihoods to choose which parameter set
to keep. New parameter sets with a higher likelihood than the old set
replaced old ones, and those with a lower likelihood replaced current
ones with a probability equal to the log-likelihood ratio. Given that we
only used the likelihoods in the decision, we used an uninformative
prior. Once we decided whether the new parameter set would replace the
old one, we ran the model again to sample the likelihood distribution of
the parameter set. We then started the cycle again by proposing a new
set of parameters and repeated the process for \(1e^5\) steps. We ran 5
independent chains, discarded the first \(1000\) samples of each chain
as burn in, and after that, we kept 1 in every 100 samples to avoid
autocorrelation. The collection of parameters sets kept in all chains
approximates the posterior distribution. We coded the model as well as
the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R (R Core Team 2021). All codes are accessible at
\href{https://zenodo.org/badge/latestdoi/440585701}{\includegraphics{https://zenodo.org/badge/440585701.svg}}.

@@

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism cleaners
use to account for the future effects of their actions; and thus to
solve the ephemeral reward task. In the estimation of the parameter
values of the full model, which includes both chaining and penalty, the
bulk of the marginal posterior distribution of \(\eta\) which controls
the strength of penalty is around 0 (Fig. \ref{fig:post}). As for
\(\gamma\), controlling chaining, the 95\% confidence intervals also
includes zero, but the mode of the posterior is around 0.5 (Fig.
\ref{fig:post}, a). In the chaining model, where \(\eta\) is set to
zero, the distribution of \(\gamma\) shifted to higher values, zero is
no longer part of the 95\% credible interval of the parameter (Fig.
\ref{fig:post}, c). In contrast, when we look at the model with only
penalty, the posterior distribution of \(\eta\) is still centred around
zero (Fig. \ref{fig:post}, d). Thus, the analysis of the estimates of
individual parameter values in the three models only supports a strong
effect of chaining. Furthermore, the comparison of the models` fit
favours the chaining model. In panel e of figure \ref{fig:post} we show
the distribution of \(pseudo-R^2\) calculated using samples from the
posterior distributions shown before. Note, \(pseudo-R^2\) can have
negative values, which is when the log-likelihood of the model is lower
than that of a model that triggers neutral preferences. Even though the
peak of the three \(pseudo-R^2\) distributions were not very different,
the model with only chaining produced a distribution of \(pseudo-R^2\)
where more values were positive (to the right of the black line in Fig.
\ref{fig:post} e). This shows that accounting for variation in the
parameter estimates the model with chaining gives a better fit to the
data, despite having one parameter less than the full model. We have not
shown here the marginal posterior distributions of the scaling constant,
given that they do not bring biological insight. Their visualizations
can be found in the supplementary material (Fig. \ref{fig:scaConst}), as
well as the diagnostics of the MCMCs (Figs.
\ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagNeg}).

@@

\begin{longtable}[]{@{}rc@{}}
\caption{\label{tab:param} Parameter values with which the model was run
in the MCMC. \(\sigma\) refers to the amplitude of the perturbation
kernel with the subscript indicating the associated parameter. New
values were taken from a uniform distribution. \(\alpha\) refers to the
learning rate.}\tabularnewline
\toprule
Parameter & Value \\
\midrule
\endfirsthead
\toprule
Parameter & Value \\
\midrule
\endhead
Learning rounds & 10000 \\
Reward value & 1 \\
\(\alpha\) & 0.05 \\
\(\sigma_{\gamma}\) & 0.3 \\
\(\sigma_{\eta}\) & 4 \\
\(\sigma_{Sca.Const.}\) & 300 \\
Number of chains & 5 \\
Chain length & \(1^5\) \\
\bottomrule
\end{longtable}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage{float}
\usepackage{graphicx}
\usepackage{booktabs}
\linespread{2}
\usepackage{lineno}
\linenumbers
\usepackage{setspace}\doublespacing
\newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords---}} #1}

@@

To capture with the model the relationship between the ecological
variables and cleaner fish preferences for visitors, we needed to scale
the absolute population densities of cleaner fish from the empirical
data to a measure of relative abundances that captures client visitation
patterns. This is because, in the model, relative abundances of clients
define not only the probability of residents and visitors but also how
often the cleaning station is empty (e.g.~there are no clients to be
cleaned). The frequency with which clients visit the station is another
variable influencing the station occupancy, which in nature may vary
among different client species depending on their ectoparasite loads. We
do not have field estimates for species-specific parasite loads,
especially not as a function of the site. In order to control for these
aspects, we computed a measure of relative cleaner fish abundance for
each reef site relative to absolute abundances and multiplied it by a
scaling constant that changes the range of the variable. Specifically,
cleaner fish relative abundance is equal to
\(\frac{\epsilon C_{abs}}{\epsilon C_{abs}+R_{abs}+V_{abs}}\). Where
\(\epsilon\) is the scaling constant; \(C_{abs}\) is the cleaner fish
absolute density; \(R_{abs}\) is the residents absolute density and
\(V_{abs}\) is the visitor client absolute density. This scaling
constant is meant to capture variation in the market conditions driven
partly by cleaner fish abundance. We fitted the value of the scaling
constant (\(\epsilon\)) as part of the statistical inference. As for the
visitor and resident abundances, we computed a relative measure with
respect to the total client abundance, and weighted that by the
re-scaled relative absence of cleaners
(\(1-\text{relative cleaner abundance}\)). Thus, all three measures of
relative abundance sum up to 1, and can be used in the model as a proxy
for the probability of having different options in the cleaning station.
Note that we introduced the scaling constant to control for variation
that is not captured by the model; its parameter distribution does not
offer biological insights.

Once we calculated the relative abundances, we obtained predictions from
the model for each one of the locations and ran the Markov Chain. We
started the chain with random values for the parameters of interest,
then ran the computational model once for every reef site using as input
the ecological explanatory variables. The model outputs the probability
of choosing the visitor for each location \(p_i\), where \(i\) is the
index for the 12 locations. Assuming a binomial distribution, the
probability that each of the cleaner fish 20 choices in the ephemeral
reward task was generated by the model is given by
\(\binom{n}{k_ij}p^{k_{ij}}_i (1-p_i)^{n-k_{ij}}\); where \(k_{ij}\) is
the number of times that cleaner fish \(j\) in reef site \(i\) chose the
visitor over the resident; and \(n=20\) (due to our choice of using 20
choices per cleaner fish). By taking the natural logarithm and summing
over all the cleaner fish and reef sites, we obtained the log-likelihood
of the data given the model and parameters. We then proposed a new set
of parameters drawn from a uniform distribution centered around the old
parameter set. The amplitude of the uniform distribution used for each
parameter can be found in Table \ref{tab:param}. Subsequently, we ran
the model and calculated the likelihood with the new parameter set. We
then used the ratio of the two likelihoods to choose which parameter set
to keep. New parameter sets with a higher likelihood than the old set
replaced old ones, and those with a lower likelihood replaced current
ones with a probability equal to the log-likelihood ratio. Given that we
only used the likelihoods in the decision, we used an uninformative
prior. Once we decided whether the new parameter set would replace the
old one, we ran the model again to sample the likelihood distribution of
the parameter set. We then started the cycle again by proposing a new
set of parameters and repeated the process for \(1e^5\) steps. We ran 5
independent chains, discarded the first \(1000\) samples of each chain
as burn in, and after that, we kept 1 in every 100 samples to avoid
autocorrelation. The collection of parameters sets kept in all chains
approximates the posterior distribution. We coded the model as well as
the fitting algorithm in \emph{c++}, the diagnostics and visualization
in R (R Core Team 2021). All codes are accessible at
\url{https://zenodo.org/badge/latestdoi/440585701}

@@

Estimation of parameter values for the three models (full model,
chaining and penalty), support chaining as the only mechanism cleaners
use to account for the future effects of their actions; and thus to
solve the ephemeral reward task. In the estimation of the parameter
values of the full model, which includes both chaining and penalty, the
bulk of the marginal posterior distribution of \(\eta\) which controls
the strength of penalty is around 0 (Fig. \ref{fig:post}). As for
\(\gamma\), controlling chaining, the 95\% confidence intervals also
includes zero, but the mode of the posterior is around 0.5 (Fig.
\ref{fig:post}, a). In the chaining model, where \(\eta\) is set to
zero, the distribution of \(\gamma\) shifted to higher values, zero is
no longer part of the 95\% credible interval of the parameter (Fig.
\ref{fig:post}, c). In contrast, when we look at the model with only
penalty, the posterior distribution of \(\eta\) is still centered around
zero (Fig. \ref{fig:post}, d). Thus, the analysis of the estimates of
individual parameter values in the three models only supports a strong
effect of chaining. Furthermore, the comparison of the models' fit
favors the chaining model. In panel e of figure \ref{fig:post} we show
the distribution of \(pseudo-R^2\) calculated using samples from the
posterior distributions shown before. Note, \(pseudo-R^2\) can have
negative values, which is when the log-likelihood of the model is lower
than that of a model that triggers neutral preferences. Even though the
peak of the three \(pseudo-R^2\) distributions were not very different,
the model with only chaining produced a distribution of \(pseudo-R^2\)
where more values were positive (to the right of the black line in Fig.
\ref{fig:post} e). This shows that accounting for variation in the
parameter estimates the model with chaining gives a better fit to the
data, despite having one parameter less than the full model. We have not
shown here the marginal posterior distributions of the scaling constant,
given that they do not bring biological insight. Their visualizations
can be found in the supplementary material (Fig. \ref{fig:scaConst}), as
well as the diagnostics of the MCMCs (Figs.
\ref{fig:diagnosticsfull},\ref{fig:diaggam},\ref{fig:diagNeg}).

@@

\begin{longtable}[]{@{}rc@{}}
\caption{\label{tab:param} Parameter values with which the model was run
in the MCMC. \(\sigma\) refers to the amplitude of the perturbation
kernel with the subscript indicating the associated parameter. New
values were taken from a uniform distribution. \(\alpha\) refers to the
learning rate.}\tabularnewline
\toprule()
Parameter & Value \\
\midrule()
\endfirsthead
\toprule()
Parameter & Value \\
\midrule()
\endhead
Learning rounds & 10000 \\
Reward value & 1 \\
\(\alpha\) & 0.05 \\
\(\sigma_{\gamma}\) & 0.3 \\
\(\sigma_{\eta}\) & 4 \\
\(\sigma_{Sca.Const.}\) & 300 \\
Number of chains & 5 \\
Chain length & \(1^5\) \\
\bottomrule()
\end{longtable}
